---
title: "Carnival_analysis"
author: "Christian Heyer"
date: "13 1 2022"
output: html_document
---

```{r, echo = F, include=F}
contrast_groups <- ifelse(exists("snakemake"),
  snakemake@wildcards[["contrast"]],
  ""
)
```

# Carnival results from deseq2 for `r contrast_groups`

Here we analyze the results from CARNIVAL analysis from each comparison group-

```{r setup, include=FALSE}
library(CARNIVAL)
library(readr)
library(piano)
library(dplyr)
library(ggplot2)
library(tibble)
library(tidyr)
library(dplyr)
library(scales)
library(plyr)
library(GSEABase)
library(network)
library(reshape2)
library(cowplot)
library(pheatmap)
library(ggraph)
library(tidygraph)
library(snowfall)
```


## Carnival Network

CARNIVAL analyses was run following generally https://github.com/saezlab/transcriptutorial 

Here we load up a network for each Comparison group and analyze it herein.

```{r snakemakeparams}
if (exists("snakemake")) {
  carnival_path <- snakemake@input[["carnival_obj"]]
  tt_basepath <- snakemake@params[["tutorial_source_path"]]
  comp_group <- snakemake@wildcards[["contrast"]]
  out_file <- snakemake@output[[1]]
  organism <- snakemake@config[["organism"]]
} else {
  yaml <- yaml::read_yaml("../../../configs/VascAge_cre_config.yaml")
  comp_group <- names(yaml$diffexp$contrasts$condition)[1]
  base_path <- yaml$dirs$BASE_ANALYSIS_DIR
  carnival_path <- file.path(
    base_path, "results/inversecarnival",
    glue::glue("{comp_group}_carnival_res.RDS.gz")
  )

  tt_basepath <- "../transcriptutorial"
  out_file <- ""
}
source(file.path(tt_basepath, "support_enrichment.r"))
source(file.path(tt_basepath, "support_networks.r"))

process_carnival <- function(carnival_res) {
  carnival_res$weightedSIF <- carnival_res$weightedSIF %>% dplyr::filter(Weight != 0)
  carnival_res
}
msig_h <- msigdbr::msigdbr(organism, "C2", )
pathways <- loadGSC(
  file = dplyr::select(
    msig_h,
    c("gene_symbol", "gs_name")
  ),
  addInfo = dplyr::select(
    msig_h,
    c("gs_name", "gs_description")
  )
)
```


### GSEA

The following gene set enrichment using a hyper geometric test to analyze:
- all Connections inferred by carnival
- Background: All nodes found in the prior knowledge network

All Networks need to interpreted in terms of the differential expression analysis.
Node activität relates to the change in expression between the comparison groups. 
Positive values <- Group1, negative values Group2.

```{r GSEA analysis}
carnival_result <- readRDS(carnival_path)
if (length(carnival_result) == 0) {
  write("CARNIVAL FAILED", file = out_file)
  quit(save = "no", status = 0)
}

carnival_result <- process_carnival(carnival_result)
nodes_carnival <- extractCARNIVALnodes(carnival_result)

sig_pathways <- runGSAhyper(
  genes = nodes_carnival$sucesses,
  universe = nodes_carnival$bg, gsc = pathways
)
sig_pathways_df <- as.data.frame(sig_pathways$resTab) %>%
  tibble::rownames_to_column(var = "pathway")

pathways_select <- sig_pathways_df %>%
  dplyr::select(pathway, `p-value`, `Adjusted p-value`) %>%
  dplyr::filter(`Adjusted p-value` <= 0.05) %>% # Signficance value fixed for now
  dplyr::rename(pvalue = `p-value`, AdjPvalu = `Adjusted p-value`) %>%
  dplyr::mutate(pathway = as.factor(pathway))


pathways_select <- data.frame(t(apply(pathways_select, 1, function(r) {
  aux <- unlist(strsplit(sub("_", ";", r["pathway"]), ";"))
  r["pathway"] <- gsub("_", " ", aux[2])
  return(c(r, "source" = aux[1]))
})))

if (ncol(pathways_select) == 4) {
  colnames(pathways_select) <- c("pathway", "pvalue", "AdjPvalu", "source")
  pathways_select$AdjPvalu <- as.numeric(pathways_select$AdjPvalu)

  ggdata <- pathways_select %>%
    dplyr::slice_min(AdjPvalu, n = 25) %>%
    dplyr::filter(AdjPvalu <= 0.05) %>%
    dplyr::group_by(source) %>%
    dplyr::arrange(AdjPvalu) %>%
    dplyr::slice(1:5)

  # Visualize top results
  ggplot(ggdata, aes(
    y = reorder(pathway, -log10(AdjPvalu)),
    x = -log10(AdjPvalu)
  ), color = source) +
    facet_grid(source ~ ., scales = "free", space = "free") +
    geom_bar(stat = "identity") +
    annotation_logticks(sides = "bt") +
    theme_bw() +
    theme(
      axis.title = element_text(face = "bold", size = 12),
      axis.text.y = element_text(size = 6)
    ) +
    ylab("")
}
```

##  Carnival network plot

Carnival network plots give an overview of the network inferred by carnival.
Changes reflect the comparison based on the comparison group from this document

```{r pressure, fig.height=10, fig.width=16}
plot_carn_network <- function(carn_res) {
  ed <- carn_res$weightedSIF
  colnames(ed) <- c("from", "sign", "to", "weight")
  nod <- data.frame(union(ed$from, ed$to))

  colnames(nod) <- "nodes"
  nod$label <- nod$nodes
  joined_nod <- left_join(nod, carn_res$nodesAttributes,
    by = c("label" = "Node")
  )
  yote <- tidygraph::tbl_graph(
    nodes = joined_nod,
    edges = ed
  ) %>%
    ggraph(layout = "auto") +
    geom_node_point(aes(color = AvgAct), size = 8) + scale_color_gradient2() +
    geom_edge_link(arrow = arrow(), aes(
      edge_colour = as.factor(sign),
      edge_alpha = weight
    )) +
    theme_graph() + geom_node_text(aes(label = label), vjust = 0.4, repel = T)
  yote
}

plot_carn_network(carnival_result) + ggtitle(comp_group)
```

