---
title: "gsea Report"
author: "Christian Heyer"
date: "1 4 2022"
output: html_document
fig_width: 12
fig_height: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "png", 
                      fig.width = 12, fig.height = 12)
library(purrr)
library(clusterProfiler)
library(magrittr)
library(ggplot2)
library(org.Mm.eg.db)
library(svglite)
if (exists("snakemake")) {
  gsea_list <- snakemake@input[["gsea_result"]]
  plot_path <- snakemake@params[["plot_path"]]
  c_groups <- snakemake@params[["contrast_groups"]]
  names(gsea_list) <- names(c_groups)
} else {

  conf <- yaml::read_yaml("../../configs/workflow_unit_test.yaml")
  BASE_DIR <- conf$BASE_DATA_DIR
  gsea_list <- as.list(file.path(BASE_DIR, glue::glue("results/diffexp/{names(conf$diffexp$contrasts$condition)}.gseares.RDS")))
  c_groups <- names(conf$diffexp$contrasts)
  names(gsea_list) <- c_groups
   gsea_list <- gsea_list[1:3]
  plot_path <- glue::glue("{BASE_DIR}/reports/deseq2")
}
```

## GSEA report

This report shows an overview othe MSIG DB gene set enrichment analysis performed 
for the Vascular Aging Project. Results are shown for each sample group and for each
GSEA gene set tested. 

```{r Parse data}

create_all_gsea_plots <- function(msig_list, contrast_n, limit = 30, type, save_plots = F) {
  if (is.list(msig_list)) {
    msig_list <- msig_list[[1]]
  }
  if (nrow(msig_list) < limit) {
    sig_ids <- msig_list@result$ID
  } else {
    sig_ids <- msig_list@result$ID[1:limit]
  }
  all_plots <- purrr::map(sig_ids, function(gset_id) {
    enrichplot::gseaplot2(
      x = msig_list,
      geneSetID = gset_id,
      title = msig_list[gset_id, ]$Description
    )
  })
  names(all_plots) <- sig_ids
  if (save_plots) {
    gsea_plot_path <- glue::glue("{plot_path}/{contrast_n}_plots/{type}", recursive = T)
    dir.create(gsea_plot_path, recursive = T)
    purrr::map2(all_plots, sig_ids, function(p, n) {
      ggsave(plot = p, filename = file.path(gsea_plot_path,paste0( n, "_gseaplot.svgz")),device = svglite)
    })
  }
  all_plots
}
gsea_result <- purrr::map(gsea_list, readRDS)
```


## Gene set enrichment (GSEA MsigDB) {.tabset}


Two tab groups are shown -> The first one controls the gene set from MSIGDB that is being
enriched against and the second tab group the comparison contrast

In case no plot is shown, then no enrichments against msigdb could be found. 
First the table with enrichment results is shown, then there is a prompt for
showing the GSEA result plots for each gene set together with the genes enriched
with their scores and if they are part of the core enrichment as determined by
GSEA: 

### GSEA MsigDB Hallmarks {.tabset}

MSIG db hallmarks results

```{r plot Enrichments, results='asis', fig.width=12, fig.height=12, echo = FALSE}
build_gene_set_detail <- function(gset_res, gset_name, inkey) {
  ensembl_names <- gset_res@result[gset_name, ]$core_enrichment %>%
    stringr::str_split("/") %>%
    unlist()


  all_geneList <- gset_res@geneList %>% tibble::as_tibble()
  all_geneList[inkey] <- as.character(names(gset_res@geneList))
  all_geneList <- all_geneList %>%
    dplyr::filter(!!rlang::sym(inkey) %in% gset_res@geneSets[[gset_name]]) %>%
    mutate(core_enrichment = ifelse(!!rlang::sym(inkey) %in% ensembl_names, "YES", "NO"))
  t_table <- clusterProfiler::bitr(all_geneList %>% dplyr::pull(inkey), inkey, "SYMBOL", OrgDb = org.Mm.eg.db)

  yarp <- dplyr::inner_join(all_geneList, t_table) %>%
    dplyr::relocate(SYMBOL) %>%
    dplyr::arrange(dplyr::desc(value))
  yarp
}


create_msig_entries <- function(gsea_res, msig_type, ktype, limit = 30) {
  stopifnot(!is.null(names(gsea_res)))
  for (set_n in names(gsea_res)) {
    cat("#### ", set_n, " {.tabset} \n")

    yote <- gsea_res[[set_n]][[msig_type]]
    if (is.list(yote)) {
      yote <- yote[[1]]
    }
    yote <- setReadable(yote, org.Mm.eg.db, keyType = ktype)

    msig_res <- gsea_res[[set_n]][[msig_type]]
    if (is.list(msig_res)) {
      msig_res <- msig_res[[1]]
    }
    all_plots <- create_all_gsea_plots(msig_res,
      type = msig_type,
      contrast_n = set_n,
      limit = limit, save_plots = T
    )
    if (nrow(msig_res@result) < limit) {
      names(all_plots) <- msig_res@result$ID
    } else {
      names(all_plots) <- msig_res@result$ID[1:limit]
    }
    print(knitr::kable(yote@result[, -c(1, length(yote@result))] %>% arrange(desc(NES))))

    cat("\n")
    for (sig_gset in names(all_plots)) {
      cat("##### ", sig_gset, " \n")

      plot(all_plots[[sig_gset]])
      cat("\n")
      gene_enrich_table <- build_gene_set_detail(msig_res,
        gset_name = sig_gset,
        inkey = ktype
      )
      print(knitr::kable(gene_enrich_table))
      cat("\n")
    }
    cat("\n")

    cat("\n \n")
  }
  cat("\n")
}

create_msig_entries(gsea_result, msig_type = "msig_gsea", ktype = "ENSEMBL", limit = 30)
```


### GSEA C3 {.tabset}

C3 results 

```{r gseac3, results='asis', fig.width=12, fig.height=12, echo = FALSE}
create_msig_entries(gsea_result, msig_type = "msig_C3", ktype = "ENSEMBL")
```


### GSEA C6 {.tabset}


C6 results 

```{r gseac6, results='asis', fig.width=12, fig.height=12, echo = FALSE}
create_msig_entries(gsea_result, msig_type = "msig_C6", ktype = "ENSEMBL")
```


### GSEA Kegg {.tabset}

Kegg

```{r kegg, results='asis', fig.width=12, fig.height=12, echo = FALSE}
create_msig_entries(gsea_result, msig_type = "kegg", ktype = "ENTREZID")
```

